Determinism, Complexity, and Predictability in Computer Performance

### 2014-02-10 pecomplexity.tex
more content

\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}
\newcommand{\gcc}{{\tt 403.gcc}~}
\newcommand{\svd}{{\tt dgesdd}~}
\newcommand{\row}{{\tt row\_major}~}
\newcommand{\col}{{\tt col\_major}~}

@@

@@

$$s=\frac{1}{n-1}\sum^n_{i=2}|X_{i,obs}-X_{i,obs}|$$ is the average in-sample
forecast error for a random walk prediction $(p_j=c_{j-1})$. This error method
was introduced in \cite{MASE} as a ``generally applicable measurement of
forecast accuracy without the problems seen in the other measurements." The
major advantage of MASE is that it allows fair comparison across methods,
prediction horizons and varying signal scales. A $MASE<1$ means that the
prediction method gives, on average, smaller errors than the 1-step errors from
the random walk forecast strategy. Analogously, a $MASE>1$ means that the
prediction method did worse, on average than the 1-step errors from the random
walk forecast strategy. In Table \ref{tab:error} we provide the distribution of
MASEs for 15 runs of each of the 8 signals across the 3 prediction forecast
methods. For comparison Table \ref{tab:error} also has weighted permutation

@@

%\item Paragraph on how to interpret MASE
\begin{enumerate}

\item First paragraph: WPE is a good measure of predictability.  Figure:
best athlete MASE vs. its WPE.

@@

\item Talk about structure analysis. Just because there is forward information
transfer, does not mean that linear predictors can get at this.
\item For this show a figure of (a) ARIMA vs MASE (b) LMA vs MASE in a side by
side plot.
\item full results.  Image: MASE vs. WPE for both LMA \& ARIMA.  Points to make:
\begin{enumerate}
\item clusters are distributed differently
\item clusters are shaped differently---tight or not
\item clusters move differently between LMA and ARIMA
\item finally, the diagonal line is important. If you're below it, you could do
better.

\end{enumerate}
\end{enumerate}

@@

>>>>

\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}

\usepackage{pifont} % check and x marks
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\gcc}{{\tt 403.gcc}~}
\newcommand{\svd}{{\tt dgesdd}~}
\newcommand{\row}{{\tt row\_major}~}
\newcommand{\col}{{\tt col\_major}~}

@@

\begin{enumerate}
\item \cmark~Introduce MASE.
\item \cmark~WPE is a good measure of predictability.  Figure: best athlete MASE vs.
its WPE.
\item \cmark~Talk about structure analysis. Just because there is forward information
transfer, does not mean that linear predictors can get at this.
\item For this show a figure of (a) ARIMA vs MASE (b) LMA vs MASE in a side by
side plot.
\item full results.  Image: MASE vs. WPE for both LMA \& ARIMA.  Points to make:
\begin{enumerate}
\item \cmark~clusters are distributed differently
\item \cmark~clusters are shaped differently---tight or not
\item \cmark~clusters move differently between LMA and ARIMA
\item finally, the diagonal line is important. If you're below it, you could do
better.
\end{enumerate}
\end{enumerate}

@@

$$s=\frac{1}{n-1}\sum^n_{i=2}|X_{i,obs}-X_{i,obs}|$$ is the average in-sample
forecast error for a random walk prediction $(p_j=c_{j-1})$. This error method
was introduced in \cite{MASE} as a ``generally applicable measurement of
forecast accuracy without the problems seen in the other measurements." The
major advantage of MASE is that it allows fair comparison across methods,
prediction horizons and varying signal scales. A $\textrm{MASE}<1$ means that
the prediction method gives, on average, smaller errors than the 1-step errors
from the random walk forecast strategy. Analogously, a $\textrm{MASE}>1$ means
that the prediction method did worse, on average, than the 1-step errors from the
random walk forecast strategy. In Table \ref{tab:error} we provide the
distribution of MASEs for 15 runs of each of the 8 signals across the 3
prediction forecast methods. For comparison Table \ref{tab:error} also has
weighted permutation

@@

@@

@@

Other noteworthy features of the LMA and ARIMA results are the cluster locations
and distributions. The WPE values of each run any particular program tend to
have little variance, leading to the clusters in
Figures~\ref{fig:arima_pred_vs_wpe}, \ref{fig:lma_pred_vs_wpe} to be fairly
constrained in the $y$ direction. For most traces, the LMA and ARIMA variance is
low too, resulting is small, tight clusters. The ARIMA MASE values of the \col
traces, however, have a large variance resulting in the spread seen in
Fig~\ref{fig:arima_pred_vs_wpe}. Not only are the MASE values of that cluster
bad, in that other predictors vastly outperform it, but they are inconsistent.
Furthermore, since LMA can predict nonlinear behavior while ARIMA can not, we
see that the clusters in Fig.~\ref{fig:arima_pred_vs_wpe} are mostly further to
the right than those in Fig.~\ref{fig:lma_pred_vs_wpe}.
